---
title: "The Rcpp plugin for FLasher"
author: Finlay Scott, Iago Mosqueira - European Commission Joint Research Center
date: "`r format(Sys.time(), '%d %B, %Y')`"
documentclass: article
header-includes: 
  - \usepackage{float}
  - \usepackage{microtype}
  - \usepackage[T1]{fontenc}
  - \usepackage[bitstream-charter]{mathdesign}
output:
  pdf_document:
    toc: true
    fig_caption: yes
    number_sections: yes
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{FLasher}
tags: [FLR FLasher forecast fisheries]
abstract:
license: European Union Public Licence (EUPL) V.1.1
---

```{r, pkgs, echo=FALSE, message=FALSE}
library(Rcpp)
library(FLasher)
library(ggplotFL)
library(knitr)
opts_chunk$set(dev='png', cache=FALSE, fig.width=6, fig.height=6, tidy=TRUE, dpi=300, fig.pos='!h')
options(width=60)
```

# Introduction

With the **Rcpp** package it is possible to write, compile and call C++ code on the fly during an R session using the *cppFunction()* and *sourceCpp()* functions.
A plugin has been written that allows the C++ components of **FLasher** to be used during an R session, including access to all of the FLCpp classes (the C++ implementation of the FLR classes) and automatic differentiation (AD) functionality through access to the CppAD library.

# Using *cppFunction()* and *sourceCpp()*

Here we demonstrate how the **Rcpp** functions *cppFunction()* and *sourceCpp()* can be used

## *cppFunction()*

*cppFunction()* is used for writing functions in C++ that you want to call from R.
You write your C++ function using standard C++ types for the arguments and returned object and the automatic **Rcpp** *as<>* and *wrap* takes care of the conversion.
The C++ function is passed as a string to *cppFunction()* during the R session:

```{r}
cppFunction('
int my_add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')
```

The C++ function can then be called as if it was an R function:

```{r}
my_add(1L, 2L, 10L)
```

It is possible to use C++11 functionality, for example, using range-based loops and auto types:
To do this we need to use C++11 plugin. This function takes a vector of integers and increments each value in the vector.

```{r, warning=FALSE}
cppFunction('
std::vector<int> rbl_demo(std::vector<int> v){
    for (auto& i : v){
        i++;
    }
    return v;
    }',
    plugins="cpp11"
)
```

We can call it as if it was a normal R function.

```{r}
rbl_demo(c(1L,2L,3L))
```

## *sourceCpp()*

*sourceCpp* is for writing longer C++ scripts and can contain multiple functions and classes, not all of which need to be exposed to R.
Exposing the desired functions to R is done using the *Rcpp::attributes*.
The C++ code can either be included as a text string or written in a separate file.
Writing the code in a separate file makes it easier to manage and also your text editor will highlight the syntax correctly.
You need to include the *include* to get all the advantages of Rcpp.
Here we write the C++ code to a temporary file and source it:

```{r}
source_code <- "
    #include <Rcpp.h>
    // This function is not exposed to R
    double foo(double x){
        return 2.0 * x;
    }

    // This function is exposed to R and calls the unexposed one
    // [[Rcpp::export]]
    double call_foo(double x){
        double y = foo(x);
        return y;
    }
"
cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

```{r}
call_foo(3.5)
```

C++11 code can be included using the C++11 plugin:

```{r, warning=FALSE}

source_code <- "
    // [[Rcpp::plugins(cpp11)]]     
    #include <Rcpp.h>

    // [[Rcpp::export]]
    std::vector<double> rbl_demo2(std::vector<double> v){
        for (auto& i : v){
            i = i * 2.0;
        }
        return v;
    }
"
cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

```{r}
rbl_demo2(c(1.3, 2.6, 3.9))
```
# Using the **FLasher** plugin

Using the **FLasher** plugin means that you have access to the methods and classes in the C++ code of the **FLasher** package.
For example, you can pass in and manipulate *FLQuant*s.

To use it with *cppFunction()* you must specify it as a *depends* argument:

```{r}
cppFunction('
FLQuant calc_catches(FLQuant landings, FLQuant discards){
    FLQuant catches = landings + discards;
    return catches;
    }',
    depends="FLasher"
)
```

```{r}
data(ple4)
landings <- landings.n(ple4)[,ac(2000:2003)]
discards <- discards.n(ple4)[,ac(2000:2003)]
calc_catches(landings, discards)
```


To use with *sourceCpp()* you must add a *depends* to top of the script and *include* the **FLasher** header file:

```{r}
source_code <- "
    // [[Rcpp::depends(FLasher)]]
    #include <FLasher.h>

    // [[Rcpp::export]]
    FLQuant calc_catches2(FLQuant landings, FLQuant discards){
        FLQuant catches = landings + discards;
        return catches;
    }
"
cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

```{r, "demo"}
calc_catches2(landings, discards)
```

# Using automatic differentiation

As well as providing access to the FLCppad classes and methods, the plugin allows the AD library CppAD to be accessed.
Unfortunately, at the moment, the interface to is a bit clunky.

Here we write C++ code that returns the value and the gradient of the *banana* function (see help page for *optim* for more information on the banana function).
We can pass the exposed gradient function to R's optim functions.
There is also an exposed function that returns the Hessian.

The function *func()* can be rewritten to be *any* function that you want that gradient etc. for.
The rest of the code remains the same (it would be good to have this other code in the package but it is not possible at the moment).

```{r, "adexample"}
source_code <- "
    // [[Rcpp::depends(FLasher)]]
    #include <FLasher.h>

    // This is the function want to solve - the banana function
    // It is templated because we need versions of it that deal with
    // types double (for normal evaluation) and adouble (for AD evaluation) 
    template <typename T>
    std::vector<T> func(std::vector<T> params){
        std::vector<T> res(1, 0.0);
        res[0] = 100 * pow((params[1] - params[0] * params[0]), 2.0) + pow((1 - params[0]), 2.0);
        return res;
    }

    // Evaluates the function
    // [[Rcpp::export]]
    std::vector<double> eval_function(std::vector<double> params){
        return func(params);
    }

    // Uses CppAD magic to get the gradient
    // [[Rcpp::export]]
    std::vector<double> eval_gradient(std::vector<double> params){
        std::vector<adouble> x(params.begin(), params.end());
        CppAD::Independent(x);
        std::vector<adouble> res = func(x);
        CppAD::ADFun<double> fun(x, res);
        return fun.Jacobian(params);
    }

    // Uses CppAD magic to get the Hessian
    // [[Rcpp::export]]
    std::vector<double> eval_hessian(std::vector<double> params, unsigned int var = 0){
        std::vector<adouble> x(params.begin(), params.end());
        CppAD::Independent(x);
        std::vector<adouble> res = func(x);
        CppAD::ADFun<double> fun(x, res);
        return fun.Hessian(params, var);
    }
"

cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

We can test this by solving the function in R with *optim()* using an approximate gradient, the exact gradient function and the AD gradient.

```{r}
# Rosenbrock Banana function
fr <- function(x) {   
    100 * (x[2] - x[1] * x[1])^2 + (1 - x[1])^2
}
# The exact gradient of the banana function
grr <- function(x) { ## Gradient of 'fr'
    x1 <- x[1]
    x2 <- x[2]
    c(-400 * x[1] * (x[2] - x[1] * x[1]) - 2 * (1 - x[1]),
        200 *      (x[2] - x[1] * x[1]))
}

# Solving with:
# Approximate gradient
res1 <- optim(c(-1.2,1), fr, method = "BFGS")
# Exact gradient
res2 <- optim(c(-1.2,1), fr, grr, method = "BFGS")
# AD gradient defined above
res3 <- optim(c(-1.2,1), eval_function, eval_gradient, method = "BFGS")

res1[c("par", "value", "counts")]
res2[c("par", "value", "counts")]
res3[c("par", "value", "counts")]
```
The version with the AD gradient is exactly the same as the version with the exact gradient function.

We can also get the Hessian:

```{r}
# Estimated
optimHess(res1$par, fr)
# Estimated using the the gradient function
optimHess(res2$par, fr, grr)
# Calculated using the AD function
eval_hessian(res3$par)
```

The above C++ code can be used to provide the gradients and Hessians for any functions.
All the user needs to do is write their own *func()* function (with the same arguments).


# Function pointers

It is possible to get hold of the pointer to a C++ function and pass that as an argument to another function.
This mechanism could be used to pass user-defined functions to routines contained in a package like **FLasher**.
The process is a bit fiddly and requires setting a *typedef* for a function, the user function that follows the typedef, a function to return a pointer the user function and a function that evaluates a function pointer.


UNFORTUNATELY WHEN THIS IS INCLUDED IN THE VIGNETTE, IT BREAKS ON BUILD

IT WORKS WHEN RENDERED INDEPENDENTLY THOUGH

NO IDEA

This example is based on http://gallery.rcpp.org/articles/passing-cpp-function-pointers/.

<!--

```{r, fp_source}
source_code <- "
    #include <Rcpp.h>

    // Make a definition for the expected function arguments
    typedef std::vector<double> (*funcPtr)(std::vector<double> x); 

    // User defined function that has the same arguments as our typedef
    // Not necessarily exported
    std::vector<double> myfunc(std::vector<double> x){
        std::vector<double> res(1, 0.0);
        res[0] = 100 * pow((x[1] - x[0] * x[0]), 2.0) + pow((1 - x[0]), 2.0);
        return res;
    }

    // Function to return a pointer to the user function
    // Note that the name of the function it points to is hardwired and must match our function above
    // [[Rcpp::export]]
    Rcpp::XPtr <funcPtr> get_func_ptr(){
        return (Rcpp::XPtr<funcPtr> (new funcPtr(&myfunc)));
    }

    // Exported function to evaluate a function passed in by address
    // [[Rcpp::export]]
    std::vector<double> call_func_ptr(SEXP xpsexp, std::vector<double> x) {
        Rcpp::XPtr<funcPtr> xpfun(xpsexp);
        funcPtr fun = *xpfun;
        std::vector<double> y = fun(x);
        return (y);
    }
"

cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

We first get the pointer to the user function, then we pass the pointer to the function that evaluates the user function.

```{r, eval_fp}
# Get the pointer
myfunc_ptr <- get_func_ptr()
myfunc_ptr
# Evaluate it
call_func_ptr(myfunc_ptr, c(-1.2, 1))
```

Why is this interesting? A function similar to *call_func_ptr()* could be included a package.
This means that some C++ routines could be contained in the **FLasher** package that use user defined functions.

A function, *eval_user_function()* is in the **FLasher** to test this.

```{r, eval_user_source}
source_code <- "
    #include <Rcpp.h>

    typedef std::vector<double> (*funcPtr)(std::vector<double> x); 

    std::vector<double> myfunc(std::vector<double> x){
        std::vector<double> res(1, 0.0);
        res[0] = 100 * pow((x[1] - x[0] * x[0]), 2.0) + pow((1 - x[0]), 2.0);
        return res;
    }

    // [[Rcpp::export]]
    Rcpp::XPtr <funcPtr> get_func_ptr(){
        return (Rcpp::XPtr<funcPtr> (new funcPtr(&myfunc)));
    }
"

cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

We can evaluate the user defined function in the same way using the existing function in **FLasher**.
However, it causes the vignette to crash so it is not run.
Weirdly, it runs outside of the vignette. I don't understand and give up.

Booooooo!

```{r, eval_user FLasher, eval=FALSE}
myfunc_ptr <- get_func_ptr()
# Evaluate it using the exposed function in FLasher
eval_user_function(myfunc_ptr, c(-1.2, 1))
```

The real interest in this lies in getting gradients etc for user defined functions, similar to the example above but without having to write the equivalent of the *eval_gradient()* function.

An attempt at this has been added to **FLasher**.
Unfortunately, although the user function is called, the returned gradients are always 0, like there is no connection between dependent and independent variables.
(See the commented out code as the bottom of src/test_solver.cpp).

The failing example, similar to the above example but using type *adouble* is below.
It compiles OK, but the gradients are not correct.

```{r, eval_grad_source}
#source_code <- "
#    // [[Rcpp::depends(FLasher)]]
#    #include <FLasher.h>
#
#    typedef std::vector<adouble> (*funcADPtr)(std::vector<adouble> x); 
#
#    std::vector<adouble> myADfunc(std::vector<adouble> x){
#        std::vector<adouble> res(1, 0.0);
#        res[0] = 100 * pow((x[1] - x[0] * x[0]), 2.0) + pow((1 - x[0]), 2.0);
#        return res;
#    }
#
#    // [[Rcpp::export]]
#    Rcpp::XPtr <funcADPtr> get_func_adptr(){
#        return (Rcpp::XPtr<funcADPtr> (new funcADPtr(&myADfunc)));
#    }
#"
#
#cat(source_code, file=paste(tempdir(),"test.cpp", sep="/"))
#sourceCpp(file=paste(tempdir(),"test.cpp", sep="/"))
```

We try to tape and evaluate the function, but the gradient is always 0.0.


```{r, eval_grad_source_test}
#myfunc_ptr <- get_func_adptr()
## Evaluate it
#eval_user_gradient(myfunc_ptr, c(-1.2, 1))
## Something really fishy is going on because this call is allowed (and wrong)
#eval_user_function(myfunc_ptr, c(-1.2, 1))
```

-->
