---
title: "Introduction to FLasher"
author: Finlay Scott, Iago Mosqueira - European Commission Joint Research Center
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
urlcolor: blue
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{FLasher}
tags: [FLR FLasher forecast fisheries]
abstract:
license: European Union Public Licence (EUPL) V.1.1
---

```{r, pkgs, echo=FALSE, message=FALSE, warnings=FALSE}
library(FLasher)
library(ggplotFL)
library(knitr)
opts_chunk$set(dev='png', cache=FALSE, fig.width=4, fig.height=4, tidy=TRUE, dpi=72)
options(width=60)
```

# Introduction

__FLasher__ is an FLR package for performing stochastic projections of fish populations and fisheries through time. It is a sequel to an older package __FLash__ that has been widely used for projecting **FLStock** objects.
It is anticpated that __FLasher__ eventually replaces __FLash__. The interfaces are almost the same so moving existing scripts from using __FLash__ to __FLasher__ should be straightforward.

The main differences betwen __FLasher__ and __FLash__ are:

* As well as performing projections on single **FLStock** objects, __FLasher__ can perform projections on a pair of **FLBiol(s)** and **FLFishery(ies)** objects to simulate mixed fishery interactions.
* __FLasher__ can operate on a seasonal as well an annual timestep.
* __FLasher__ is built using the __CppAD__ library, whereas __FLash__ was build using __ADOLC__ (this should be invisible to the user).

# Basic principles

A projection is controlled by a **fwdControl** object. This determines how long the projection runs for, and what targets the projection attempts to hit along the way.

When projecting with **FLBiol(s)** and **FLFishery(ies)** objects, the fishing efforts of the **FLFishery(ies)** in each timestep are found to hit the target.

Projecting with an **FLStock** is the equivalent of projecting with a single **FLBiol** and **FLFishery** (this is what happens internally).
When an **FLStock** is being projected, the Fmultiplier in each step is found that hits the target (the assumption is made that effort and F are linearly related).

# Moving from __FLash__ to __FLasher__

If you are moving from using __FLash__ to __FLasher__ there are a few basic things to note.

The first thing to note is argument names to the *fwd()* function.
The projection is still run using the *fwd()* function. The **FLStock** argument can still be unnamed if it is passed in as the first argument. However, the **fwdControl** argument must now be named as *control*.  Additionally, the name of the *sr.residuals* argument is now simply called *residuals*

```{r, eval=FALSE}
# Will not work
out <- fwd(stock, control, sr=srrbits, sr.residuals=residuals)
# Will work
out <- fwd(stock, control=control, sr=srrbits, residuals=residuals)
```

The second thing is that when making the **fwdControl** object the names of the some of the columns has changed.
The *quantity* column is now called *quant*.
The *val* column is now called *value*.

Other differences will be added as we remember them.

# Where to start

See the tutorials at the FLR website:

* The [tutorial](http://www.flr-project.org/doc/Running_Medium_Term_Forecasts_with_FLasher.html) on running medium term projections with an **FLStock**.
* The [tutorial](http://www.flr-project.org/doc/Setting_Stock_Recruitment_in_FLasher_Projections.html) on setting stock recruitment for __FLasher__ projections.
* The [tutorial](http://www.flr-project.org/doc/Mixed_Fisheries_Projections_with_FLasher.html) on performing mixed fisheries projections with __FLasher__.
* The [tutorial](http://www.flr-project.org/doc/Using_the_FLasher_Plugin.html) vignette on using the __FLasher__ plugin for accessing Automatic Differentiation capabilities in R and other nerdy things.

More details of the internal workings of __FLasher__ can be found in the vignette *FLasher reference manual*.

